


\subsection{Latex sizes}

XXX REMOVE ME XXX

\begin{itemize}

\item colummwidth: \the\columnwidth
\item textwidth:  \the\textwidth
\item parskip: \the\parskip
\item abovecaption: \the\abovecaptionskip
\item belowcaption: \the\belowcaptionskip


\end{itemize}

\subsection{Misc. to cite}

Stuff to cite (that is not otherwise cited) ... that we should either incorporate or drop:



\begin{enumerate}

\oldcite{Corey OS}{DBLP:conf/osdi/Boyd-WickizerCCMKMPSWDZZ08}
\oldcite{Thekah -- early User-level networking}{DBLP:conf/sigcomm/ThekkathNML93}

\oldcite{Libra JVM}{DBLP:conf/vee/AmmonsABSGKKRHW07}
\oldcite{Scheduler activations}{DBLP:journals/tocs/AndersonBLL92}
%\oldcite{Atikoglu - Workload analysis fo large-scale key-value store}{Atikoglu:2012:WAL}
\oldcite{Basu - Efficient virutal memory for big memory servers}{DBLP:conf/isca/BasuGCHS13}
\oldcite{RadixVM}{DBLP:conf/eurosys/ClementsKZ13}
%\oldcite{Waldspurger - CPU balloning}{DBLP:conf/osdi/Waldspurger02}
%\oldcite{Drepper - cost of virtualization}{DBLP:journals/queue/Drepper08}
\oldcite{Thialfi}{DBLP:conf/sosp/AdyaCMP11}
%\oldcite{Cavium Octeon}{cavium-octeon}
\oldcite{Fan 2013 Mem Cache 3}{fan2013memc3}
%\oldcite{Case against userlevel networking}{magoutis2004case}


\ana{Could cite Harchol-Balter, Schroeder, Bansal, and Agrawal. Size-based Scheduling to Improve Web Perfomrance.}


\ana{Should cite Mogul and Ramakrishnan. Eliminating Receive Livelock in an Interrupt-Driven Kernel.}


\ana{From OSDI program committee, might consider citing: Emin Gun Sirer and Nick Feadmster.}

\end{enumerate}

\subsection{Dangling from related work}


\paragraph{Adaptive, batched run to completion}

\todo TODO

\paragraph{Coherence-free, flow-consistent processing}

\todo Affinity Accept, mTCP, ...



\paragraph{API design.}


\paragraph{Hardware specialization - Dataplanes.}

\paragraph{Operating system architecture.}

\paragraph{Domain-specific and library operating systems.}



\input{tbl-pingpong}
