
\section{Introduction}

\todo Data Center trends ... webscale applications

\todo Current pain points.  Facebook relies on UDP~\cite{nishtala2013scaling}; Google's long tail~\cite{DBLP:journals/cacm/DeanB13}

\todo The C10M challenge~\cite{theC10Mproblem}  scale servers to massive connection counts, deliver 10\microsecond~latency (mean), allowing for an extra 10\microsecond~of jitter; saturate multiple 10GbE links.   Conventional wisdom is that it must be userlevel.

\todo Energy-propoortionality challenge~\cite{DBLP:journals/computer/BarrosoH07}

\todo Introducing IX.

\todo Primary contributions. 

\begin{itemize}
\item first networking stack that meets the C10M challenge
\item despite the conventional wisdom, actually is a protected, OS-based networking stack
\item energy-proportional because of dynamic spinning cores.
\end{itemize}



\section{Related Work}

\todo need a table that compares us with related work (!)

\todo Improving operating system kernels to scale

\todo Limitations of the POSIX socket API

\todo User-level networking

\todo Protocol offload engines

\todo Energy-proportionality.   Have a graph that shows the energy-proportionality of a dynamic core approach.


\section{Design Principles}

\todo Separation of control and data planes

\todo Coherency-free operations  (ala multikernel~\cite{DBLP:conf/sosp/BaumannBDHIPRSS09})

\todo Dynamic spinning cores : re-define pthread abstraction but use it in a different way

\todo Asynchronous channels (like Megapipe)

\todo Abstraction layers: libevent, libpcap


\section{Implementation}

\todo System Archictecture (Dune~\cite{belay2012dune})

\todo Driver layer

\todo TODO (figures, ...)

\section{Evaluation}


\todo Compare apples-to-apples with Megapipe, and mTCP in terms of microbenchmarks.

\todo Baseline comparison of state-of-the art systems include:  Linux (some recent version, with SO\_REUSEPORT); Megapipe (if possible), mTCP (if possible). 

\todo Microbenchmark: short TCP transactions (echo server, as defined in megapipe).   Goal is to beat mTCP (and therefore all others) hands down.

\todo Benchmark: memcached - compare with FB results (?)

\todo Benchmark: lightttpd -- used by Affinity-Accept and mTCP.  

\todo ngnx: an actually used webserver.


\section{Discussion}
\section{Conclusion}




